Detecting changes in dispersion in COVID-19 incidence time series using a negative binomial model

### 2023-05-11 noise_covid_v2.Rmd
remove excess code

---
title: Detecting changes in dispersion in COVID-19 incidence time series using a negative binomial model
author: "Rachael Aber"
header-includes: \usepackage{setspace}\doublespacing
output:
  html_document:
    df_print: paged
  pdf_document: default
geometry: margin=2.54cm
fontsize: 12pt
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, 
                      fig.align = "center", fig.width = 5, fig.height = 5)
set.seed(12345)
devtools::load_all()
source("R/simulation.R")
```

@@

Approximately two months of data on either side of November 24th, 2020 was used in the analysis. This step was done to include a set number of "waves" of incidence in the model fit to each side of the hypothesized change. However, we acknowledge that there is a trade-off between including few waves and increasing the power of the proposed test by including more time series data.

```{r, echo=FALSE, fig.cap="COVID-19 county-level incidence in New York surrounding Thanksgiving 2020."}
# USA case/population size data to work with

data <- read.csv("data/confirmed_and_death_1209.csv")
data$DateTime <- as.Date(data$DateTime)

# NY dataframe for Thanksgiving analysis
data <- data[data$DateTime >= "2020-10-26" & data$DateTime <= "2020-12-24", ]
inds <- seq(as.Date("2020-10-26"), as.Date("2020-12-24"), by = "day")
days <- weekdays(inds)
inc <- aggregate(DailyNewConfirmedCases ~ countyFIPS, data, FUN = list)
inc$populations <- tapply(X = data$population, INDEX = data$countyFIPS, FUN = mean)
inc$stateFIPS <- as.numeric(substr(inc$countyFIPS, 1, 2))

# all_zero <- c()
# for (i in 1:dim(inc)[1]){
#   if (all(unlist(inc$DailyNewConfirmedCases[i]) == 0)){
#     print(i)
#     all_zero <- c(all_zero, i)
#   }
# }
# inc <- inc[-all_zero,] 
ny_df_t <- inc[inc$stateFIPS == 36,]
fl_df_t <- inc[inc$stateFIPS == 12,]
ny_fl_df_t <- rbind(ny_df_t, fl_df_t)
```

@@

scaled_cases <- map(ny_df_t$DailyNewConfirmedCases, ~ as.vector(scale(.)))

ny_df_t$ch_pts <- map(scaled_cases, ~ cpts(cpt.mean(., class = TRUE, penalty = "MBIC", pen.value = 0, method = "PELT", param.estimates = TRUE)))

which(sapply(ny_df_t$ch_pts, length) == 0)

@@

ny_df_t$ch_pts2 <- map(ny_df_t$DailyNewConfirmedCases, ~ cpts(cpt.meanvar(., class = TRUE, penalty = "MBIC", pen.value = 30, method = "PELT", test.stat = "Poisson", param.estimates = TRUE, minseglen = 2))) 
ny_df_t$n_ch_pts2 <- map(ny_df_t$ch_pts2, ~length(unlist(.)))

@@

par(mfrow=c(1,2))
# By visual inspection 
plot(ny_df_t$DailyNewConfirmedCases[[1]] ~ inds, type="l", ylim=c(0, 5000), xaxt="n", ylab="Daily Incidence", xlab="Date")
for (i in 2:dim(ny_df_t)[1]){
  lines(ny_df_t$DailyNewConfirmedCases[[i]] ~ inds, col=sample(rainbow(10)))
}
abline(v=inds[30], lwd=2)
axis(1, inds, inds, cex.axis = .7)

@@

@@

time <- 1:length(spline_mod.1.nb$mu)
plot(time, Y[1:(length(Y)/2)], xlab="Day Index (Segment 1)", ylab="Incidence") 
lines(time, spline_mod.1.pois$fitted.values, col="red")
lines(time, spline_mod.1.nb$mu, col="blue")
legend(10, 10, legend=c("Poisson", "NB"),
       col=c("red", "blue"), lty=1)
```

>>>>

---
title: Detecting changes in dispersion in COVID-19 incidence time series using a negative binomial model
author: "Rachael Aber"
header-includes: \usepackage{setspace}\doublespacing
output:
  html_document:
    df_print: paged
  pdf_document: default
geometry: margin=2.54cm
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, 
                      fig.align = "center", fig.width = 5, fig.height = 5)
```

@@

Approximately one month of data on either side of November 24th, 2020 was used in the analysis. This step was done to include a set number of "waves" of incidence in the model fit to each side of the hypothesized change. However, we acknowledge that there is a trade-off between including few waves and increasing the power of the proposed test by including more time series data.

@@

library(tidyverse)
library(changepoint)

filename <- "data/processed_dat.Rdata"
load(filename)
rm(filename)

scaled_cases <- map(as.data.frame(t(new_cases_subset)), ~ as.vector(scale(.)))

ch_pts <- map(scaled_cases, ~ cpt.mean(., class = TRUE, penalty = "MBIC", pen.value = 0, method = "PELT", param.estimates = TRUE))

@@

ny_df_t$ch_pts2 <- map(as.data.frame(new_cases_subset), ~ cpts(cpt.meanvar(., class = TRUE, penalty = "MBIC", pen.value = 30, method = "PELT", test.stat = "Poisson", param.estimates = TRUE, minseglen = 2))) 
ny_df_t$n_ch_pts2 <- map(ny_df_t$ch_pts2, ~length(unlist(.)))

@@

inds <- seq(start_date, end_date, by = "day")

plot(new_cases_subset[state_index[1],], type = "l", xaxt = "n", ylab = "Daily Incidence", xlab = "Date")
for (i in 2:length(state_index)){
  lines(new_cases_subset[state_index[i],], col=sample(rainbow(10)))
}
abline(v=inds[30], lwd=2)
axis(1, inds, inds, cex.axis = .7)

@@

```{r}
source("code/squared.res_v_squared.fitted.R")
```

@@

source("code/check_nb_estimation.R")
```
